---
title: "POMP_GroupLevel_ManuscriptOutline"
author: "Madeline Peters, Aaron King, Nina Wale"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package loading and data preparation
## Load required packages

```{r packages}
library(tidyverse)
library(stringi)
library(doParallel)
library(pomp)
library(panelPomp)
library(ggpubr)
```

## Set working directory

```{r directory}
setwd("~/Documents/GitHub/bdd/nw11_hier/final_files/")
```

## Colour-blind friendly colour palette

```{r palette}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


## Read in PNAS data

```{r data, echo=FALSE}
read_csv("data.csv",
         col_types="iiinnnn"
) %>%
  mutate(
    mouseid=sprintf("%02d-%02d",box,mouse),
    box=sprintf("%02d",box),
    mouse=sprintf("%02d",mouse),
    paba=as.factor(box),
    rbc_density=rbc_density/1000
  ) %>%
  mutate(
    paba=recode(
      paba,
      "01"="0.05","02"="0.005","03"="0.0005","04"="0","05"="control"
    )
  ) %>%
  select(
    day,
    Pd=ama_density,
    RBC=rbc_density,
    Ter119=ter119_density,
    CD71=cd71_density,
    mouseid,
    paba,box,mouse
  ) %>%
  arrange(mouseid,day) %>%
  mutate(
    paba=as.character(paba),
    Ter119=ifelse(Ter119==0,NA,Ter119),
    CD71=ifelse(CD71==0,NA,CD71)
  ) %>%
  mutate(
    Eryth=(1-CD71/Ter119)*RBC,
    Retic=CD71/Ter119*RBC
  ) -> flow
head(flow)
```

## Obtain estimates for beta and dose

```{r beta, echo=FALSE}
flow %>%
  filter(day<=4) %>%
  lm(log(Pd)~box:day+mouseid-1,data=.) -> fit2

coef(fit2) %>% 
  bind_rows() %>%
  gather(var,val) %>%
  mutate(
    var=stri_replace_all_regex(var,"mouseid(\\d{2})-(\\d{2})","dose[$1-$2]"),
    var=stri_replace_all_regex(var,"box(\\d{2}):day","Beta[$1]"),
    val=exp(val)
  ) -> theta1

expand.grid(
  box=sprintf("%02d",1:5),
  mouse=sprintf("%02d",1:3)
) %>%
  mutate(
    mouseid=paste0(box,"-",mouse),
    betavar=sprintf("Beta[%s]",box),
    dosevar=sprintf("dose[%s]",mouseid)
  ) %>%
  left_join(theta1,by=c("betavar"="var")) %>%
  rename(Beta=val) %>%
  left_join(theta1,by=c("dosevar"="var")) %>%
  rename(dose=val) %>%
  select(-betavar,-dosevar) %>%
  arrange(box,mouse) %>%
  mutate(
    Beta=coalesce(Beta,0),
    dose=coalesce(dose,0)
  ) -> theta

registerDoParallel()

theta %>%
  mutate(
    sigmaPd = 2, 
    sigmaRBC = 0.1, 
    sigmaRetic = 0.3, 
    sigmaW = 1, 
    sigmaN = 0.5, 
    sigmaR = 0.5,
    E_0 = 8.0e6, 
    R_0 = 3e5, 
    W_0 = 8.8e4, 
    N_0 = 8e5
  ) -> theta
head(theta)
```

## Create object pos with pomp object for each mouse  

```{r pomps}
foreach (m = iter(theta,"row"),.inorder=TRUE,.combine=c) %dopar% {
  
  flow %>%
    filter(mouseid==m$mouseid) %>%
    select(day,Pd,RBC,Retic) %>%
    mutate(Retic=if_else(day %in% c(0,14),NA_real_,Retic)) %>%
    pomp(
      params=select(m,-mouseid,-box,-mouse) %>% unlist(),
      times="day",
      t0=0,
      rmeasure=Csnippet("
  Retic = rlnorm(log(1+R),sigmaRetic)-1;
  RBC = rlnorm(log(1+E+R),sigmaRBC)-1;
  Pd = rlnorm(log(1+K),sigmaPd)-1;"),
      dmeasure=Csnippet("
  double l1, l2, l3;
  l1 = (R_FINITE(Retic)) ? dlnorm(1+Retic,log(1+R),sigmaRetic,1) : 0;
  l2 = (R_FINITE(RBC)) ? dlnorm(1+RBC,log(1+E+R),sigmaRBC,1) : 0;
  l3 = (R_FINITE(Pd) && Pd>0) ? dlnorm(1+Pd,log(1+K),sigmaPd,1) : 0;
  lik = (give_log) ? l1+l2+l3 : exp(l1+l2+l3);"),
      rprocess=discrete_time(
        step.fun=Csnippet("
  double Mold = M;
  M = Beta*K*exp(-(W+N)/(R+E));
  E = (R+E)*exp(-(Mold+N)/(R+E));
  N = rlnorm(log(N),sigmaN);
  W = rlnorm(log(W),sigmaW);
  R = rlnorm(log(R),sigmaR);
  K = (R+E>0) ? (R+E)*(1-exp(-M/(R+E))): 0;
  "),
        delta.t=1
      ),
      partrans=parameter_trans(
        log=c("sigmaW","sigmaR","sigmaN",
              "sigmaPd","sigmaRBC","sigmaRetic",
              "N_0","W_0","E_0","R_0")
      ),
      rinit=Csnippet("
  E = E_0;
  R = R_0;
  N = N_0;
  W = W_0;
  M = 0;
  K = dose;"),
      statenames=c("E","R","W","N","M","K"),
      paramnames=c(
        "Beta","dose",
        "sigmaPd","sigmaRBC","sigmaRetic",
        "sigmaW","sigmaN","sigmaR",
        "E_0","R_0","W_0","N_0"
      )
    )
} %>% 
  set_names(theta$mouseid) -> pos
```

## Obtain MLE for sigmas, initial values, betas and dose

```{r MLE, echo=FALSE}
pf4name <- "m5pf4.rds"
pf4 <- readRDS(pf4name)

pf4 %>%
  select(loglik,starts_with("sigma"),ends_with("_0")) %>%
  filter(loglik==max(loglik)) -> mle

mle %>% unlist() -> p
pos %>% panelPomp(shared=p) %>% coef() %>% 
  rbind() %>% as_tibble() -> theta
theta %>%
  select(starts_with("Beta"),starts_with("dose")) %>%
  gather(variable, value) %>%
  tidyr::extract(variable, into=c("variable","mouseid"), 
                 regex="([[:alnum:]]+)\\[(.+?)\\]") %>%
  spread(variable,value) -> bdf
print(mle)
print(bdf)
```

## Load in PNAS trajectories
```{r trajectories}
sm1name <- "m5sm1.rds"
sm1 <- readRDS(sm1name)
```

## Create dataframe with weighted trajectories (grouped by mouseid first, then by box)

```{r traj_df}
sm1 |>
  as_tibble() |>
  filter(mouseid!="01-02",mouseid!="02-03") |> #remove underdosed mice
  pivot_wider(names_from=variable,values_from=value) |>
  left_join(bdf,by=c("mouseid")) |>
  separate_wider_delim(cols="mouseid",delim="-",names=c("box","mouse"),cols_remove=FALSE) |>
  group_by(mouseid) |>
  mutate(
    SM=exp(-M/(R+E)),
    SN=exp(-N/(R+E)),
    SW=exp(-W/(R+E)),
    Qps=(1-SM)*SW*SN,
    Qpn=N/(N+W)*(1-SM)*(1-SW*SN),
    Qpw=W/(N+W)*(1-SM)*(1-SW*SN),
    Qun=SM*(1-SN),
    Qus=SM*SN,
    lambda_r=Beta*R/M*Qps,
    lambda_e=Beta*E/M*Qps,
    lambda_n=Beta*(R+E)/M*Qpn,
    lambda_w=Beta*(R+E)/M*Qpw,
    lambda_u=Beta-lambda_r-lambda_e-lambda_n-lambda_w,
    lambda_u_w_ratio=lambda_u/lambda_w,
    lambda_u_n_ratio=lambda_u/lambda_n,
    lambda_u_wn_ratio=lambda_u/(lambda_w+lambda_n),
    loss=(R+E)*(1-Qus),
    rbc=E+R,
    varN=N/rbc,
    perRetic=R/rbc,
    lik=exp(loglik-max(loglik))
  ) |>
  ungroup() |>
  select(-loglik,-Beta,-dose) |>
  gather(variable,value,-rep,-time,-mouse,-box,-lik,-mouseid) |>
  filter(is.finite(value)) %>%
  group_by(box,time,variable) |>
  dplyr::reframe(
    value=pomp::wquant(x=value,probs=c(0.05,0.5,0.95),weights=lik),
    name=c("lo","med","hi")
  ) |>
  ungroup() |>
  pivot_wider() -> group_traj

group_traj$pABA <- factor(group_traj$box,levels=c("05","04","03","02","01"),
                          labels=c("Uninfected","0%","0.0005%","0.005%","0.05%"))
```

## Remove estimates for W for control mice
```{r removeControlW}
group_traj <- group_traj |> dplyr::slice(-which(group_traj$box=="05"&group_traj$variable=="W"))
```

# Data visualisation

## Does infected alter reticulocyte supply as a function of time?
```{r ReticvsTime_UninfInf}
group_traj_UninfInf<-group_traj |>
  filter(pABA%in%c("Uninfected","0%")) |>
  mutate(status=case_match(pABA,
                          "Uninfected"~"Uninfected",
                          "0%"~"Infected"))

group_traj_UninfInf |>
  filter(variable=="R") |>
  ggplot()+
  geom_ribbon(aes(x=time,ymin=lo,ymax=hi,fill=status),alpha=0.2)+
  geom_line(aes(x=time,y=med,col=status),linewidth=2)+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), 
                  labels=trans_format('log10',math_format(10^.x)))+ 
  scale_colour_manual(values=cbPalette[c(2,1)],name=NULL)+
    scale_fill_manual(values=cbPalette[c(2,1)],name=NULL)+
  xlab("Day post-infection")+
  ylab("Reticulocyte supply (density per microlitre)")+
  theme_bw()+
  theme(
    strip.background=element_blank(),
    panel.grid = element_blank(),
    legend.position=c(0.2,0.8),
    axis.title=element_text(size=15),
    axis.text=element_text(size=13),
    legend.text=element_text(size=14))
```

## Is reticulocyte supply a linear function of erythrocyte density?
```{r ReticvsTime_UninfInf}
group_traj_UninfInf |>
  filter(variable%in%c("E","R")) |>
  select(-lo,-hi) |>
  pivot_wider(names_from="variable",values_from="med") |>
  ggplot()+
  geom_path(aes(x=E,y=R,col=status),linewidth=2)+
  geom_text(aes(x=E,y=R,label=time))+
  scale_colour_manual(values=cbPalette[c(2,1)],name=NULL)+
  xlab("Erythrocytes (density per microlitre)")+
  ylab("Reticulocyte supply (density per microlitre)")+
  theme_bw()+
  theme(
    strip.background=element_blank(),
    panel.grid = element_blank(),
    legend.position=c(0.8,0.8),
    axis.title=element_text(size=15),
    axis.text=element_text(size=13),
    legend.text=element_text(size=14))
```

## Can reticulocyte supply during infection be divided into separate phases?
```{r ReticvsTime_UninfInf}
group_traj_UninfInf |>
  filter(variable=="R") |>
  ggplot()+
  geom_ribbon(aes(x=time,ymin=lo,ymax=hi,fill=status),alpha=0.2)+
  geom_line(aes(x=time,y=med,col=status),linewidth=2)+
  geom_bracket(
      xmin = 0.1, xmax = 6.9, y.position = 6.5,
      label = "Manipulation\nresponse", tip.length = c(0.01, 0.01),hjust=0.5
    )+
  geom_bracket(
      xmin = 7.1, xmax = 9.9, y.position = 6.5,
      label = "Stress\nerythropoiesis", tip.length = c(0.01, 0.01),hjust=0.5
    )+
  geom_bracket(
      xmin = 10.1, xmax = 29.9, y.position = 6.5,
      label = "Reticulocyte supply ~ erythrocyte density", tip.length = c(0.01, 0.01),hjust=0.5
    )+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), 
                  labels=trans_format('log10',math_format(10^.x)))+ 
  scale_colour_manual(values=cbPalette[c(2,1)],name=NULL)+
    scale_fill_manual(values=cbPalette[c(2,1)],name=NULL)+
  xlab("Day post-infection")+
  ylab("Reticulocyte supply (density per microlitre)")+
  theme_bw()+
  theme(
    strip.background=element_blank(),
    panel.grid = element_blank(),
    legend.position=c(0.2,0.7),
    axis.title=element_text(size=15),
    axis.text=element_text(size=13),
    legend.text=element_text(size=14))
```